02-antecedentes

### 2024-06-13 02-antecedentes.qmd

While meritocratic preferences refer to a justification of distribution based on merit criteria (effort and talent), the perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019]. This perception can vary greatly depending on individual experiences, social, economic, and cultural background. Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work. In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2021]. Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger market justice preferences, as individual achievement would be seen as rewarded and social policies as less necessary.

Some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2021], and how larger inequality affects meritocratic beliefs [@morris_representing_2022]. A larger justification of meritocratic distribution have been related to less support for redistributive compensation systems [@frank_performance_2015]. Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018]. @almas_cutthroat_2020 found that in the US the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit). @barr_effect_2020 found that in unequal societies, the highly educated accept inequality more than the low educated. Using data for 41 countries from the International Social Survey Programme (ISSP),@garcia-sanchez_attitudes_2020 found that the perceived size of the income gap correlated positively with support for progressive taxation, an association that was weaker among those who endorsed meritocratic and equal opportunity beliefs. In the same line, experimental research by @durante_preferences_2009 showed that subjects support less redistribution when the initial distribution is determined according to task performance.

>>>>

While meritocratic preferences refer to a justification of distribution based on merit criteria (effort and talent), the perception of meritocracy refers to how individuals view and understand the concept of meritocracy in their own society [@duru-bellat_who_2012; @castillo_meritocracia_2019]. This perception can vary greatly depending on individual experiences, social, economic, and cultural background. Some people may see meritocracy as a fair and just system that allows anyone to succeed based on their abilities and hard work. In contrast, others may view it as a myth or a cover for existing power dynamics and inequality, serving to maintain and even reinforce inequality [@lampert_meritocratic_2013; @mijs_paradox_2019]. Based in this last perspective, we argue that individuals with a higher perception of meritocracy will show a larger market justice preferences, as individual achievement would be seen as rewarded and social policies as less necessary.

Some studies have analyzed how those with greater privileges believe more in meritocracy (Reynolds & Chan 2014), how greater economic inequality increases meritocratic beliefs [@mijs_paradox_2019], and how larger inequality affects meritocratic beliefs [@morris_representing_2022]. A larger justification of meritocratic distribution have been related to less support for redistributive compensation systems [@frank_performance_2015]. Conversely, individuals tend to support redistribution when they believe that the disadvantaged lack the opportunities to succeed [@evans_strong_2018]. @almas_cutthroat_2020 found that in the US the highly educated accept inequality significantly more than the less educated because they perceive inequality as justifiable owing to differences in productivity (i.e., merit). @barr_effect_2020 found that in unequal societies, the highly educated accept inequality more than the low educated. Using data for 41 countries from the International Social Survey Programme (ISSP),@garcia-sanchez_attitudes_2020 found that the perceived size of the income gap correlated positively with support for progressive taxation, an association that was weaker among those who endorsed meritocratic and equal opportunity beliefs. In the same line, experimental research by @durante_preferences_2009 showed that subjects support less redistribution when the initial distribution is determined according to task performance.
